```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Import Data Below
# read_csv()
```

```{r}
#read CSV files

HALElifeExpectancy = read_csv("HALElifeExpectancyAtBirth.csv")
AlcoholSubstanceAbuse = read_csv("alcoholSubstanceAbuse.csv")
medicalDoctors = read_csv("medicalDoctors.csv")
uhcCoverage = read_csv("uhcCoverage.csv")
crudeSuicideRates = read_csv("crudeSuicideRates.csv")
roadTrafficDeaths = read_csv("roadTrafficDeaths.csv")
reproductiveAgeWomen = read_csv("reproductiveAgeWomen.csv")
adolescentBirthRate = read_csv("adolescentBirthRate.csv")
eliminateViolenceAgainstWomen = read_csv("eliminateViolenceAgainstWomen.csv")
`3070cancerChdEtc` = read_csv("30-70cancerChdEtc.csv")
airPollutionDeathRate = read_csv("airPollutionDeathRate.csv")
atLeastBasicSanitizationServices <- read_csv("atLeastBasicSanitizationServices.csv")
```

```{r}
#Aditi Question 1

reproductiveAgeWomen_Clean <- 
  reproductiveAgeWomen %>%
  select(-Indicator) %>%
  rename("Year" = "Period",
         "Married or in-union women of reproductive age who 
         have their need for family planning satisfied with modern methods (%)" = "First Tooltip")

adolescentBirthRate_Clean <- 
  adolescentBirthRate %>%
  select(-Indicator) %>%
  rename("Year" = "Period",
         "Adolescent birth rate (per 1000 women aged 15-19 years)" = "First Tooltip")

eliminateViolenceAgainstWomen_Clean <- 
  eliminateViolenceAgainstWomen %>%
  select(-Indicator, -Dim1) %>%
  pivot_wider(names_from = "Dim2" , 
              values_from = "First Tooltip") %>%
  rename("Year" = "Period") %>%
  mutate(Year = stringr::str_sub(Year, 1, 4))

```

```{r}
#Aditi Question 2

`3070cancerChdEtc_Clean` <-
  `3070cancerChdEtc` %>%
  select(-Indicator) %>%
  pivot_wider(names_from = "Dim1" , 
              values_from = "First Tooltip") %>%
  rename("3070_Both sexes" = "Both sexes", 
         "3070_Male" = "Male", 
         "3070_Female" = "Female", 
         "Year" = "Period")

airPollutionDeathRate_Clean_Cat <-
  airPollutionDeathRate %>%
  # select(-Indicator) %>%
  # pivot_wider(names_from = "Dim1" , 
  #             values_from = "First Tooltip") %>%
  # rename("3070_Both sexes" = "Both sexes", 
  #        "3070_Male" = "Male", 
  #        "3070_Female" = "Female", 
  #        "Year" = "Period")
```


# Creator: Kevin Lu

### Q1: What is the relationship between universal health coverage and crude suicide rates?

```{r}
#Kevin Question 1

uhcCoverage_Clean <- 
  uhcCoverage %>%
  select(-Indicator) %>%
  rename("Year" = "Period",
         "UHC index of essential service coverage" = "First Tooltip")

crudeSuicideRates_Clean <- 
  crudeSuicideRates %>%
  select(-Indicator) %>%
  pivot_wider(names_from = "Dim1" , 
              values_from = "First Tooltip") %>%
  rename("CSR_Both sexes" = "Both sexes", 
         "CSR_Male" = "Male", 
         "CSR_Female" = "Female", 
         "Year" = "Period")

uhc_CSR_join <- uhcCoverage_Clean %>%
  full_join(crudeSuicideRates_Clean, by = c("Location", "Year"))
kevq1 <- uhc_CSR_join %>%
  drop_na()
ggplot(kevq1, aes(x=`UHC index of essential service coverage`, y=`CSR_Both sexes`)) +
  geom_point() +
  geom_smooth(method= 'lm') +
  ggtitle("UHC Index vs Crude Suicide Rates") +
  labs(x="UHC Index of Service Coverage (SCI)", y="Crude Suicide Rates per 100,000 Population")
```

### Q2: What is the relationship between alcohol substance abuse and road traffic deaths?

```{r}
#Kevin Question 2

roadTrafficDeaths_Clean <- 
  roadTrafficDeaths %>%
  select(-Indicator) %>%
  rename("Year" = "Period",
         "Estimated road traffic death rate (per 100 000 population)" = "First Tooltip")

AlcoholSubstanceAbuse_Clean <- 
  AlcoholSubstanceAbuse %>%
  select(-Indicator) %>%
  pivot_wider(names_from = "Dim1" , 
              values_from = "First Tooltip") %>%
  rename("ASA_Both sexes" = "Both sexes", 
         "ASA_Male" = "Male", 
         "ASA_Female" = "Female", 
         "Year" = "Period")

kevq2 <- AlcoholSubstanceAbuse_Clean %>%
  filter(Year == 2015) %>%
  inner_join(roadTrafficDeaths_Clean, by = c("Location"))
  
ggplot(kevq2, aes(x= `ASA_Both sexes`, y=`Estimated road traffic death rate (per 100 000 population)`)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle("Alcohol Substance Abuse vs Road Traffic Deaths") +
  labs(x="Total Alcohol Consumption per Capita (15+)", y="Estimated Road Traffic Death Rate per 100,000 Population")

#roadTrafficDeaths only contains values for 2016 - medicalDoctors is the only dataset with 2016 in the joined dataset
```

# Interpreter: Aditi Chiney

### Q1: WRITE_QUESTION_HERE

```{r}
#
```

### Q2: WRITE_QUESTION_HERE

```{r}
#
```

# Orator: Jessica Barta

### QUESTIONS 1 AND 2: WHICH COUNTRIES INCREASED THEIR OVERALL HEALTHCARE THE MOST? WHICH COUNTRIES DECREASED THEIR OVERALL HEALTHCARE THE MOST? 

```{r}
uhc_coverage_clean <- 
  uhcCoverage %>% 
  pivot_wider(names_from = 'Indicator', values_from = 'First Tooltip')

uhc_coverage_clean

life_expectancy_at_birth_clean <-
    HALElifeExpectancy %>% 
  pivot_wider(names_from = 'Indicator', values_from = 'First Tooltip')
life_expectancy_at_birth_clean

cancer_clean <-
  `3070cancerChdEtc` %>%
  pivot_wider(names_from = 'Indicator', values_from = 'First Tooltip')
cancer_clean_filtered = filter(cancer_clean, Dim1 == "Both sexes")

sanitation_clean <-
  atLeastBasicSanitizationServices %>%
  pivot_wider(names_from = 'Indicator', values_from = 'First Tooltip')
sanitation_clean
```
```{r}
ggplot(data = uhc_coverage_clean, aes(x = Period, y = `UHC index of essential service coverage`, group = Location)) +
  geom_line() + facet_wrap(~Location)

#there are spaces in the country names that are preventing R from reading the locations correctly. Replace all spaces with underscores.

uhc_coverage_clean_locations = str_replace_all(uhc_coverage_clean$Location, c(" " = "_"))
uhc_coverage_clean_locations
#now try graph again 

uhc_coverage_clean_with_underscores = mutate(uhc_coverage_clean, 'Location' = uhc_coverage_clean_locations)
uhc_coverage_clean_with_underscores

#the visualization is too crowded. Split up the datset to make it cleaner
```
```{r}
uhc_coverage_clean_with_underscores
```

```{r}
dataset1 =uhc_coverage_clean_with_underscores[1:50,]
dataset2 = uhc_coverage_clean_with_underscores[51:100,]
dataset3 = uhc_coverage_clean_with_underscores[101:150,]
dataset4 = uhc_coverage_clean_with_underscores[151:200,]
dataset5 = uhc_coverage_clean_with_underscores[201: 250,]
dataset6 = uhc_coverage_clean_with_underscores[250: 300,]
dataset7 = uhc_coverage_clean_with_underscores[301: 366,]
ggplot(data = dataset2, aes(x = Period, y = `UHC index of essential service coverage`, group = Location, colour = Location)) +
  geom_line() + ggtitle("Change in healthcare coverage/time")

uhc_coverage_clean_with_underscores_2 = pivot_wider(uhc_coverage_clean_with_underscores, names_from = 'Period', values_from = 'UHC index of essential service coverage') %>%
  mutate('difference' = `2017` - `2015`)

largest_changes = filter(uhc_coverage_clean_with_underscores, Location %in% c("Ireland", "Nepal", "Croatia"))

ggplot(data = largest_changes, aes(x = Period, y = `UHC index of essential service coverage`, group = Location, colour = Location)) +
  geom_line() + ggtitle("Countries with greatest change in healthcare coverage/time")

#Therefore, Croatia increased the most in terms of universal health care and Ireland and Nepal decreased the most. 

#Croatia goes from a uhs index of 60 to 71 from 2015 to 2017.

#use this code to show Croatia's change over time, because it is the most dramatic, compared to 49 other countries: ggplot(data = dataset1, aes(x = Period, y = `UHC index of essential service coverage`, group = Location, colour = Location)) + geom_line() 

#use this code to show the changes over time of healthcare coverage for the 3 countries where it changed the most: ggplot(data = largest_changes, aes(x = Period, y = `UHC index of essential service coverage`, group = Location, colour = Location)) + geom_line() 
```
# Deliverer: JULIA MENGXUAN YU

### Question 1: What’s the relationship between health expectancy and alcohol substance abuse in groups of males and females? Is there a significant difference between these two groups?
```{r}
#Julia Question 1
HALElifeExpectancy_JY <- 
  HALElifeExpectancy %>%
  select(-Indicator)  %>%
  rename("HE"="First Tooltip", 
         "Year" = "Period")
medicalDoctors_JY <- 
  medicalDoctors %>%
  select(-Indicator) %>%
  rename("Year" = "Period",
         "Medical_Doctors" = "First Tooltip")
AlcoholSubstanceAbuse_JY <- 
  AlcoholSubstanceAbuse %>%
  select(-Indicator)  %>%
  rename("asa"="First Tooltip", 
         "Year" = "Period")
HALE_MD_ASA <-HALElifeExpectancy_JY %>%
  full_join(medicalDoctors_JY, by = c("Location", "Year")) %>%
  full_join(AlcoholSubstanceAbuse_JY, by = c("Location", "Year"))
HALE_MD_ASA_new<- na.omit(HALE_MD_ASA) %>%
  filter(Dim1.x==Dim1.y)%>%
  select(-Dim1.y)%>%
  filter(Dim1.x!="Both sexes")
HALE_MD_ASA_new

ggplot(HALE_MD_ASA_new,aes(x=asa, y=HE,color=Dim1.x)) +
  geom_point(alpha=0.3) +
  geom_smooth(method = 'lm') +
  ggtitle("Alcohol Substance Abuse vs Health Expectancy") +
  labs(x="Total Alcohol Consumption per Capita (15+)", y="Health Expectancy")
```

### Question 2:How can the variable medical doctor influence health expectancy and alcohol substance abuse? What is the result of the first question after accounting for the influence of medical doctor?
```{r}
ggplot(HALE_MD_ASA_new,aes(x= Medical_Doctors, y=HE,color=Dim1.x)) +
  geom_point(alpha=0.5)+
  geom_smooth() +
  ggtitle("Medical Doctors vs Health Life Expectancy") +
  labs(x="Medical doctors (per 10,000)", y="Healthy life expectancy (HALE) at birth (years)")
ggplot(HALE_MD_ASA_new,aes(x= Medical_Doctors, y=asa,color=Dim1.x)) +
  geom_point(alpha=0.5)+
  geom_smooth() +
  ggtitle("Medical Doctors vs Alcohol Substance Abuse") +
  labs(x="Medical doctors (per 10,000)", y="Total Alcohol Consumption per Capita (15+)")
```


# Follow-up Questions

### New Questions Based Off Initial Investigation

- Q1: Does the relationship between UHC and crude suicide rates change when gender is a factor?
- Q2: What is the distribution of medical doctors look like? Is the distribution of medical doctors related to the transition in the chart"Medical Doctors vs Alcohol Substance Abuse"?
- Q3: WHAT IS THE DISTRIBUTION OF CHANGES IN THE HEALTHCARE INDEX? HOW MUCH DOES IT TEND TO VARY?
- Q4: WRITE_QUESTION_HERE

### Investigation of Follow-up Questions

GIVE WHAT 2 QUESTIONS YOU ATTEMPTED TO INVESTIGATE FURTHER IN COMPLETE SENTENCES (*Example:* Our group decided to investigate Q2 and Q4 in further detail.)

SHOW AT LEAST 2 TABLES OR FIGURES BELOW THAT EXPLORE ANSWERS FOR THE QUESTIONS YOU ARE INVESTIGATING FURTHER.

```{r}
#Kevin Follow-Up Question 1
#Does the relationship between UHC and crude suicide rates change when gender is a factor?
kevFollowUp <- crudeSuicideRates %>%
  select(-Indicator) %>%
  rename("Sex" = "Dim1", 
         "CSR" = "First Tooltip", 
         "Year" = "Period") %>% 
  filter(Sex != 'Both sexes') %>%
  full_join(uhcCoverage_Clean, by = c("Location", "Year")) %>%
  drop_na()
ggplot(kevFollowUp, aes(x=`UHC index of essential service coverage`, y=`CSR`, color=Sex)) +
  geom_point() +
  geom_smooth(method= 'lm') +
  ggtitle("UHC Index vs Crude Suicide Rates Between Males and Females") +
  labs(x="UHC Index of Service Coverage (SCI)", y="Crude Suicide Rates per 100,000 Population")
```
```{r}
#Julia Follow-up question
ggplot(medicalDoctors_JY) +
  geom_histogram(aes(x=Medical_Doctors))
ggplot(medicalDoctors_JY, aes(x = Medical_Doctors)) + 
  geom_boxplot()
summary(medicalDoctors_JY$Medical_Doctors)
```

```{r}
#Jessica Follow question1
ggplot(data = uhc_coverage_clean_with_underscores_2, mapping = aes(x = difference)) + 
  geom_boxplot()+ coord_flip() + ggtitle("Distribution of Changes in Healthcare Coverage")
```

# Summary

GIVE A 2 PARAGRAPH SUMMARY. 

PARAGRAPH 1 SHOULD DESCRIBE WHAT YOU LEARNED ABOUT YOUR DATA FROM INVESTIGATING THE INITIAL QUESTIONS. DID YOU FIND ANYTHING UNUSUAL IN YOUR DATA? DID ANYTHING SURPRISE YOU? WHICH OF THE INITIAL QUESTIONS WERE HELPFUL IN LEADING YOU TO MORE QUESTIONS?

-Kevin Lu:
-The data showed surprising results for the initial questions. I observed a positive correlation between universal health coverage and crude suicide rates, which I had expected to be a negative correlation. I also observed a negative correlation between alcohol consumption and road traffic deaths, which I had expected to be a positive correlation. 
-Julia Yu: 
-My data and plots showed me surprising results for the initial questions. I had assumed that the increase of alchohol abuse would lead to a negative influence on life expectancy. The woman group showed a stronger positive relationships. However, the first plot showed a positive relationship. So I questioned if the increasing of medical doctors would affect these two variables. As medical doctors increases life expectancy tend to be stable, while for alcohol abuse, the pattern is quite ambiguous.
-Jessica Barta
-Croatia increased healthcare coverage the most (+11) while Ireland and Nepal decreased their coverage the most (-3). Most countries didn't see much change in their coverage between 2015 and 2017, because the standard deviation was low, about 1.7.

PARAGRAPH 2 SHOULD SUMMARIZE WHAT YOU LEARNED FROM INVESTIGATING THE FOLLOW-UP QUESTIONS. WHY ARE THESE FOLLOW-UP QUESTIONS INTERESTING FOR INVESTIGATION? DESCRIBE THE TABLES/FIGURES YOU USED TO EXPLORE ANSWERS TO THESE FOLLOW-UP QUESTIONS? WHAT DID YOU LEARN FROM THE TABLES/FIGURES REGARDING THE FOLLOW-UP QUESTIONS YOU PROPOSED?
-Kevin Lu:
-The follow up question explored the relationship between universal health coverage and crude suicide rates while considering gender as a factor. The new figure showed that the positive correlation observed in the initial question is almost entirely due to a strong positive correlation for males, while there was little to no change for females based on universal health coverage. The figure also provides insight that female suicide rates are much lower than male suicide rates overall.
-Jessica Barta
-Most countries didn't see much change in their coverage between 2015 and 2017, based on the low standard deviation, about 1.7, and boxplot showing little variation. There were only a couple outliers noted above. The lack of variation brings up questions about how healthcare coverage was measured. Should the WHO have used a more sensitive measurement that would pick up more variation? It seems unlikely that nearly all the countries surveyed changed very little, or the same amount, over time.
-Julia Yu
-Most average medical doctors values are between 0 and 30. And I found the mean is about 20.685. In the chart "Medical Doctors vs Alcohol Substance Abuse", the transtion occurred at around x=20 (average medical doctors). Therefore, as the average medical doctors approximately reach to the mean, the life expectancy tend to be stable.













